---
title: "Comparing ASMR to CMD"
author: "Hardy Griesbauer"
date: "10/12/2019"
output: powerpoint_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,fig.width = 10,fig.height = 6)

library(tidyverse)
library(ggpubr)
library(forestDroughtTool)

# Load dataset
load("../data/cmdASMR.RData")

```

### Methods
- Generate historical monthly ASMR values using asmrCalc() function in forestDroughtTool package for several climate stations
- Compared to ClimateBC-derived CMD values using linear regression
- Compared over different time periods (full record vs 1961-1990 normal period) and over seasons

### Climate stations and corresponding BGC units
```{r echo=FALSE}
dat %>% 
  filter(stn_name!="LYTTON 2")  %>% # Remove this station because data too sparse
  filter(stn_name!="SMITHERS")  %>% # Remove this station because data too sparse
  mutate(bgc=replace(bgc,stn_name=="FORT ST JAMES","SBSdw3")) %>% # not sure why this was missing
  group_by(stn_name) %>% 
  summarise(bgc=first(bgc),
            Start=min(Year),
            End=max(Year)) %>% 
  mutate(Length=End-Start+1) %>% 
  knitr::kable()

```

### Soil Data used for asmrCalc
```{r echo=FALSE}
knitr::kable(soilsData)

```

### Annual CMD vs Annual ASMR
```{r include=FALSE}
X<-
  dat %>% 
  group_by(stn_name,bgc,Year) %>% # group by station and year to create annual summaries
  summarise(CMD=sum(CMD),
            xeric.ASMR=mean(S1.ASMR),
            subxeric.ASMR=mean(S2.ASMR),
            submesic.ASMR=mean(S3.ASMR),
            mesic.ASMR=mean(S4.ASMR),
            subhygric.ASMR=mean(S5.ASMR))

## Define plotting function
plotASMR<-function(BGC,site,x.pos=300,y.pos=0.8) {
  
  X %>% 
  filter(bgc==BGC) %>% 
  ggscatter(x = "CMD", y = paste(site,"ASMR",sep="."), add = "reg.line",
          add.params = list(color = "red"))+
          
 stat_cor(
   aes(label = paste(..rr.label.., ..p.label.., sep = "~`,`~")),
  label.x = x.pos,
  label.y=y.pos)

}


```

### SBSdw3 - Fort St James
```{r echo=FALSE}

bgcName="SBSdw3"

p1<-plotASMR(BGC=bgcName,site="xeric",x.pos=120,y.pos=0.65)
p2<-plotASMR(BGC=bgcName,site="subxeric",x.pos=120,y.pos=0.65)
p3<-plotASMR(BGC=bgcName,site="submesic",x.pos=120,y.pos=0.75)
p4<-plotASMR(BGC=bgcName,site="mesic",x.pos=120,y.pos=0.75)

ggarrange(p1,p2,p3,p4,nrow=2,ncol=2)


```

### PPxh3 - Grand Forks

```{r echo=FALSE}

bgcName="PPxh3"

p1<-plotASMR(BGC=bgcName,site="xeric",x.pos=400,y.pos=0.53)
p2<-plotASMR(BGC=bgcName,site="subxeric",x.pos=400,y.pos=0.58)
p3<-plotASMR(BGC=bgcName,site="submesic",x.pos=400,y.pos=0.62)
p4<-plotASMR(BGC=bgcName,site="mesic",x.pos=400,y.pos=0.68)

ggarrange(p1,p2,p3,p4,nrow=2,ncol=2)


```


### MSdk2 - Golden

```{r echo=FALSE}

bgcName="MSdk2"

p1<-plotASMR(BGC=bgcName,site="xeric",x.pos=300,y.pos=0.45)
p2<-plotASMR(BGC=bgcName,site="subxeric",x.pos=300,y.pos=0.5)
p3<-plotASMR(BGC=bgcName,site="submesic",x.pos=300,y.pos=0.53)
p4<-plotASMR(BGC=bgcName,site="mesic",x.pos=300,y.pos=0.58)

ggarrange(p1,p2,p3,p4,nrow=2,ncol=2)


```

### SBSmk1 - Prince George

```{r echo=FALSE}

bgcName="SBSmk1"

p1<-plotASMR(BGC=bgcName,site="xeric",x.pos=80,y.pos=0.7)
p2<-plotASMR(BGC=bgcName,site="subxeric",x.pos=80,y.pos=0.8)
p3<-plotASMR(BGC=bgcName,site="submesic",x.pos=80,y.pos=0.8)
p4<-plotASMR(BGC=bgcName,site="mesic",x.pos=80,y.pos=0.8)

ggarrange(p1,p2,p3,p4,nrow=2,ncol=2)

```

### IDFxh1 - Lumby

```{r echo=FALSE, fig.cap="Note only 19 years of data"}

bgcName="IDFxh1"

p1<-plotASMR(BGC=bgcName,site="xeric",x.pos=280,y.pos=0.6)
p2<-plotASMR(BGC=bgcName,site="subxeric",x.pos=280,y.pos=0.65)
p3<-plotASMR(BGC=bgcName,site="submesic",x.pos=280,y.pos=0.7)
p4<-plotASMR(BGC=bgcName,site="mesic",x.pos=280,y.pos=0.75)

ggarrange(p1,p2,p3,p4,nrow=2,ncol=2)

```


### Thoughts from annual variables

- Correlations between both models are significant in all cases, and strong in 4 of the BGC units shown.  
- Correlations are highest in hot/dry climates.
- Linear regression is not totally appropriate here, because ASMR values are bounded by 1.  If we pursue this further, should use beta regression to handle bounded data.
- Outlier years might be worth examining further

### Growing season ASMR vs growing season CMD
Next, look at relationship between CMD and ASMR over growing season (May to August)
```{r echo=FALSE}
X<-
  dat %>% 
  filter(month>=5 & month<=8) %>% # Filter for growing season
  group_by(stn_name,bgc,Year) %>% # group by station and year to create annual summaries
  summarise(CMD=sum(CMD),
            xeric.ASMR=mean(S1.ASMR),
            subxeric.ASMR=mean(S2.ASMR),
            submesic.ASMR=mean(S3.ASMR),
            mesic.ASMR=mean(S4.ASMR),
            subhygric.ASMR=mean(S5.ASMR))

```


### SBSdw3 - Fort St James - growing season
```{r echo=FALSE}

bgcName="SBSdw3"

p1<-plotASMR(BGC=bgcName,site="xeric",x.pos=120,y.pos=0.3)
p2<-plotASMR(BGC=bgcName,site="subxeric",x.pos=120,y.pos=0.4)
p3<-plotASMR(BGC=bgcName,site="submesic",x.pos=120,y.pos=0.5)
p4<-plotASMR(BGC=bgcName,site="mesic",x.pos=120,y.pos=0.6)

ggarrange(p1,p2,p3,p4,nrow=2,ncol=2)


```

### PPxh3 - Grand Forks

```{r echo=FALSE}

bgcName="PPxh3"

p1<-plotASMR(BGC=bgcName,site="xeric",x.pos=260,y.pos=0.25)
p2<-plotASMR(BGC=bgcName,site="subxeric",x.pos=260,y.pos=0.3)
p3<-plotASMR(BGC=bgcName,site="submesic",x.pos=260,y.pos=0.3)
p4<-plotASMR(BGC=bgcName,site="mesic",x.pos=260,y.pos=0.4)

ggarrange(p1,p2,p3,p4,nrow=2,ncol=2)


```

### Thoughts from growing season variables

- Correlations are stronger between growing season compared to annual;
- Correlations are again highest in hot/dry climates;
- Outlier years might be worth examining further

### Comparing slopes between soil moisture regimes
PPxh3 - Grand Forks

```{r echo=FALSE}

bgcName="PPxh3"




```


### Next steps
- Replicate with more stations?
- Look at 1961-1990 vs recent periods? Do the models agree over both time periods?
- If further analysis confirms that both models are in strong agreement for most units, then we can develop a model that predicts the regression slope and coefficient, and apply that to CMD values to estimate ASMR across soil moisture classes.