---
title: "Comparing ASMR to CMD"
author: "Hardy Griesbauer"
date: "10/12/2019"
output: powerpoint_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,fig.width = 10,fig.height = 6)

library(tidyverse)
library(ggpubr)
library(forestDroughtTool)
library(broom)

# Load dataset
load("../data/cmdASMR.RData")

```

### Methods
- Generate historical monthly ASMR values using asmrCalc() function in forestDroughtTool package for several climate stations
- Compared to ClimateBC-derived CMD values using linear regression
- Compared over different time periods (full record vs 1961-1990 normal period) and over seasons

### Climate stations and corresponding BGC units
```{r echo=FALSE}
dat %>% 
  filter(stn_name!="LYTTON 2")  %>% # Remove this station because data too sparse
  filter(stn_name!="SMITHERS")  %>% # Remove this station because data too sparse
  mutate(bgc=replace(bgc,stn_name=="FORT ST JAMES","SBSdw3")) %>% # not sure why this was missing
  group_by(stn_name) %>% 
  summarise(bgc=first(bgc),
            Start=min(Year),
            End=max(Year)) %>% 
  mutate(Length=End-Start+1) %>% 
  knitr::kable()

```

### Soil Data used for asmrCalc
```{r echo=FALSE}
knitr::kable(soilsData)

```

### Compare regression slopes between soil moisture regimes

Idea here is to understand how the coefficient between CMD and TACA varies by soil moisture regime within a BGC unit.

I regressed AET/PET against CMD for PPxh3 and SBSdw3.  Kept x- and y-axis limits same between BGC units for comparison.

```{r include=FALSE}
X<-
  dat %>% 
  group_by(stn_name,bgc,Year) %>% # group by station and year to create annual summaries
  summarise(CMD=sum(CMD),
            xeric=mean(S1.ASMR),
            subxeric=mean(S2.ASMR),
            submesic=mean(S3.ASMR),
            mesic=mean(S4.ASMR),
            subhygric=mean(S5.ASMR)) %>% 
  pivot_longer(cols=5:9,names_to="site",values_to="ASMR") # pivot longer to make plotting easier

## Define plotting function
plotSlope<-function(BGC) {
  
  X %>% 
  filter(bgc==BGC) %>% 
  filter(site!="subhygric") %>% 
  ggscatter(x= "ASMR",y="CMD",color="site",shape="site", 
            add="reg.line",
            title=BGC)+
    xlim(0.5,1)+
  ylim(100,700)
  
}


```

### PPxh3
```{r echo=FALSE, caption="Can see that slopes are essentially same between SMR"}
plotSlope("PPxh3")
```


### SBSdw3
```{r echo=FALSE, caption="Can see that slopes are essentially same between SMR"}
plotSlope("SBSdw3")
```

### Observations from the previous two slides
- Slopes are almost identical (but not completely) within a BGC unit, and differ between BGC units.  
- If we treat slopes as equal within BGC unit and among soil moisture classes, then we can model CMD differences simply using any AET/PET value, bounded by AET/PET range used. Probably a more elegant way to do this, but this approach works for now.

### CMD adjustments suggested for PPxh3, calculated using mean AET/PET ratio of 0.7


```{r echo=FALSE}

  BGC="PPxh3"

  fitCMD=c(
  
  X %>% 
    filter(bgc==BGC) %>% 
    filter(site=="mesic") %>% # remove subhygric site for now
    lm(CMD~ASMR,data=.) %>% 
    predict(object=.,newdata=data.frame(ASMR=c(0.7))),


  X %>% 
    filter(bgc==BGC) %>% 
    filter(site=="submesic") %>% # remove subhygric site for now
    lm(CMD~ASMR,data=.) %>% 
    predict(object=.,newdata=data.frame(ASMR=c(0.7))) ,


  X %>% 
    filter(bgc==BGC) %>% 
    filter(site=="subxeric") %>% # remove subhygric site for now
    lm(CMD~ASMR,data=.) %>% 
    predict(object=.,newdata=data.frame(ASMR=c(0.7))),
   

  X %>% 
    filter(bgc==BGC) %>% 
    filter(site=="xeric") %>% # remove subhygric site for now
    lm(CMD~ASMR,data=.) %>% 
    predict(object=.,newdata=data.frame(ASMR=c(0.7)))
  )  
 
cmdAdj<-data.frame(SMR=c("mesic","submesic","subxeric","xeric"),CMD.adjust=
  cumsum(c(0,abs(diff(fitCMD)))))

rownames(cmdAdj)=cmdAdj$SMR

knitr::kable(cmdAdj)


```


### CMD adjustments applied to PPxh3 data
Now, we can adjust CMD values and then plot to see if regression lines align properly.

```{r echo=FALSE}


X.adj<-
  rbind(
  filter(X,site=="mesic"),
  filter(X,site=="submesic") %>% mutate(CMD=CMD+cmdAdj["submesic",2]),
   filter(X,site=="subxeric") %>% mutate(CMD=CMD+cmdAdj["subxeric",2]),
   filter(X,site=="xeric") %>% mutate(CMD=CMD+cmdAdj["xeric",2])
) %>% 
  filter(bgc==BGC) %>% 
  arrange(Year)





```

### CMD adjustments applied to PPxh3 data
```{r echo=FALSE} 

  X.adj %>% 
  ggscatter(x= "ASMR",y="CMD",color="site",shape="site", 
            add="reg.line",
            title=BGC)
  

```


### CMD adjustments applied to PPxh3 data

Now the regression slopes align properly.  You can see the slight difference in slopes.  

### Next steps
- Need to double check soils data and confirm this is appropriate
- Use regression coefficients and slopes for each BGC/site combination to determine the CMD shift required between BGC units.  

```{r include=FALSE}
# Define function to extract coefficients and intercepts from model

extrCoef<-function(BGC) {
  
  X %>% 
    filter(bgc==BGC) %>% 
    filter(site!="subhygric") %>% # remove subhygric site for now
    group_by(site) %>% 
    do(model=lm(CMD~ASMR,data=.)) %>% 
    broom::tidy(model)
 
}

# Example
knitr::kable(extrCoef("PPxh3"))

```


